{\bf Landslide: Concurrency Bug-Finding for Pebbles}

### 2018-04-03 landslide-lecture/psu/lecture.tex
revise slides for psu suitability

\usepackage{manfnt}
\usepackage{wasysym}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{url}
\usepackage{ulem}
\usepackage{marvosym}
\usepackage{skull}
\usepackage{proof}
\usepackage{array}
\usepackage{colortbl}
\setbeamertemplate{navigation symbols}{}

\title[Landslide]{{\bf Landslide: A New Race-Finding Tool for 15-410}}
\subtitle[]{ {\em more clever than } ``\texttt{agility\_drill}'' {\em since 2011.}}
\author[Ben Blum]{Ben Blum \texttt{(bblum@andrew.cmu.edu)}}

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2018, February 16}

@@

\begin{document}
\renewcommand{\inserttotalframenumber}{39}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

\begin{frame}{Outline}
	\textbf{Theory: Seeing concurrency bugs in a new way}
	\begin{itemize}
		\item Case study (example)
		\item Tabular execution traces
		\item The execution tree
	\end{itemize}
	{\bf Research Technique: ``Systematic testing''}
	% Say: A way of enumerating all possible interleavings of a concurrent test case.
	\begin{itemize}
		\item Preemption points
		\item Challenges and feasibility
	\end{itemize}
	{\bf Tool: Landslide}
	\begin{itemize}
		\item How it works
		\item Automatically choosing preemption points
		\item User study (that's you!)
			% !!! SAY !!! : If you are struggling, this is not for you, etc.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Case Study}
	\begin{center}
		\begin{tabular}{llll}
			\fbox{
			\begin{tabular}{l}
				Consumer thread \\
				\hline
				\\
				\texttt{mutex\_lock(mx);} \\
				\\
				\texttt{if~(!work\_exists())} \\
				\texttt{~~~~cond\_wait(cvar,~mx);} \\
				\texttt{work~=~dequeue();} \\
				\\
				\texttt{mutex\_unlock(mx);} \\
				\texttt{access(work->data);} \\
			\end{tabular}
			}
			& & &
			\fbox{
			\begin{tabular}{l}
				Producer thread \\
				\hline
				\\
				\texttt{mutex\_lock(mx);} \\
				\\
				\texttt{enqueue(work);} \\
				\texttt{signal(cvar);} \\
				\\
				\texttt{mutex\_unlock(mx);} \\
				\\
				\\
			\end{tabular}
			}
		\end{tabular}
		\pause
		\linegap

		\begin{itemize}
			\item See {\bf Paradise Lost} lecture!
			\item {\tt if} vs {\tt while}: Two consumers can race to make one fail.
		\end{itemize}
	\end{center}
\end{frame}
\begin{frame}{Thread Interleavings (``good'' case)}
	\begin{center}
		\begin{tabular}{|l|l|l|}
			\hline
			\cellcolor{thread1} {\bf Thread 1} & \cellcolor{thread2} {\bf Thread 2} & \cellcolor{thread3} {\bf Thread 3} \\
			\hline
			\small \texttt{lock(mx);} & & \\
			\small \texttt{if~(!work\_exists())} & & \\
			\small \texttt{~~~~wait(cvar,~mx);} & & \\

@@

\begin{frame}{Testing}
	How can programmers be confident in the correctness of their code?
	\begin{itemize}
		\item Unit tests
		\begin{itemize}
			\item good for basic functionality, bad for concurrency
		\end{itemize}
		\item Stress tests
		\begin{itemize}
			\item state of the art in 15-410
		\end{itemize}
		\item Theorem proving
		\begin{itemize}
			\item heavy burden on the programmers
		\end{itemize}
		\item \sout{Releasing to paying customers and worrying about correctness later}
			% Say... "and I'm hoping to prevent you from finding yourself in a future where this happens!"
	\end{itemize}
	\linegap

@@

\begin{frame}{Preemption Point Example (remember this?)}
	\fbox{
	\begin{tabular}{l}
	\texttt{boolean~want[2]~=~\{~false,~false~\};}\\
	\\
	\begin{tabular}{rll}
		1 & \texttt{want[i]~=~true;} & \\
		  & & \footnotesize (preemption point A) \\
		2 & \texttt{while~(want[j])} & \\
		  & & \footnotesize (preemption point B) \\
		3 & \texttt{~~~~~~~~continue;} & \\
		  & & \footnotesize (preemption point C) \\
		4 & \texttt{//~...critical~section...} & \\
		  & & \footnotesize (preemption point D) \\
		5 & \texttt{want[i]~=~false;} & \\
	\end{tabular}
	\end{tabular}
	}
	\linegap

@@

\begin{frame}{Preemption Point Example (remember this?)}
	\fbox{
	\begin{tabular}{l}
	\texttt{boolean~want[2]~=~\{~false,~false~\};}\\
	\\
	\begin{tabular}{rll}
		1 & \texttt{want[i]~=~true;} & \\
		  & & \footnotesize (preemption point A) \\
		2 & \texttt{while~(want[j])} & \\
		  & & \footnotesize ~ \\
		3 & \texttt{~~~~~~~~continue;} & \\
		  & & \footnotesize ~ \\
		4 & \texttt{//~...critical~section...} & \\
		  & & \footnotesize ~ \\
		5 & \texttt{want[i]~=~false;} & \\
	\end{tabular}
	\end{tabular}
	}
	\linegap

	Here, only preemption point A will trigger a deadlock. {\bf ~}\\
	All other interleavings are benign.{\bf ~}

@@

\begin{frame}{About The Project}
	\textbf{About me}: Final year graduate student, advised by Garth Gibson
	\begin{itemize}
		\item TAed 15-410 for 3 semesters during undergrad
		%\item 1st graduate year was 5th year M.S., rest in Ph.D. program
		\item Landslide's publication history
		\begin{itemize}
			\item Master's thesis
			\begin{itemize}
				\item \url{http://www.contrib.andrew.cmu.edu/~bblum/thesis.pdf}
			\end{itemize}
			\item Conference paper (OOPSLA 2016)
			\begin{itemize}
				\item \url{http://www.contrib.andrew.cmu.edu/~bblum/oopsla.pdf}
			\end{itemize}
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	{\bf About Landslide}
	\begin{itemize}
		\item Simics module, which traces:
			\begin{itemize}
				\item Every instruction executed
				\item Every memory access read/written
			\end{itemize}
			% FIXME for next year -- how much slowdown is there? 20x?
		%\item Originally supported only P3s; can now test P2s fully-automated
		\item {\em Landslide} shows how your {\em Pebbles} programs may not be stable.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Big Picture: Execution Tree Exploration}
	\textbf{Backtracking}
	\begin{itemize}
		\item Each time test completes, identify a PP to replay differently
		\item Reset machine state and start over
		\item Implemented using Simics bookmarks
			\begin{itemize}
				\item {\tt set-bookmark} and {\tt skip-to}
			\end{itemize}
		\item Replay test from the beginning, with a different interleaving
	\end{itemize}
	\pause
	\linegap

	{\bf Controlling scheduling decisions}
	\begin{itemize}
		\item Tool must control all sources of nondeterminism
		\item In 15-410, just timer and keyboard interrupts
			% optional - say: "in real kernels, disk/network io interrupts"
		\item Landslide repeatedly fires timer ticks until desired thread is run.
	\end{itemize}
\end{frame}

\begin{frame}{Landslide \& You}
	\begin{center}
	\includegraphics[width=0.85\textwidth]{landslide-new-blank.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Identifying Bugs}
	Landslide can {\em definitely discover:}
	\begin{itemize}
		\item Assertion failures % Say: note: the better your asserts are..!
		\item Segfaults
		\item Deadlock
		\item Use-after-free / double-free
	\end{itemize}
	\linegap
	Landslide can {\em reasonably suspect:}
	\begin{itemize}
		\item Infinite loop (halting problem)
		\item Data race bugs
	\end{itemize}
\end{frame}

@@

	Data races are {\em not necessarily} bugs, just highly suspicious!
	\begin{itemize}
		\item Bakery alg: Is {\tt number[i]=max(number[0],number[1])+1} bad?
		%\item Is unprotected {\tt foobar\_statistics\_counter++} bad?
		\item What about unprotected {\tt next\_thread\_id++}?
		\item ``If threads interleaved the wrong way here, it {\em might} crash later.''
		\begin{itemize}
			\item Hmmm...
		\end{itemize}
	\end{itemize}
\end{frame}

@@

	State of the art tools choose a fixed set of preemption points.
	\begin{itemize}
		\item E.g., ``all thread API calls'' or ``all kernel mutex locks/unlocks''
		\item Depending on length of test, completion time is unpredictable.
		\item More often, a subset is better in terms of time/coverage. 
	\end{itemize}
	\pause
	\linegap
	Current systematic testing model is not user-friendly.
	\begin{itemize}
		\item Tool: ``I want to use these PPs, but can't predict completion time.''
		\item User: ``I have 16 CPUs and 24 hours to test my program.''
	\end{itemize}
	\linegap
	{\em Stress testing allows user to choose total run time -- can we offer this too?}
\end{frame}

@@

\begin{frame}{Test Suite}
	Landslide ships with 6 approved test cases:
	\linegap

	Standard P2 tests
	\begin{itemize}
		\item {\tt thr\_exit\_join}
		\item {\tt paraguay}
		\item {\tt rwlock\_downgrade\_read\_test}
	\end{itemize}

	New tests
	\begin{itemize}
		\item {\tt broadcast\_test}
		\item {\tt paradise\_lost} %(new this semester)
		\item {\tt mutex\_test} %(new this semester)
	\end{itemize}
\end{frame}

@@

	Sample code from {\tt cyclone}: % abbreviated
	\begin{center}
	\begin{tabular}{l}
		\texttt{for (i = 0; i < MAX\_MISBEHAVE /* 64 */; i++) \{} \\
		\texttt{~~~~misbehave(i);} \\
		\texttt{~~~~tid = thr\_create(child\_fn, i);} \\
		\texttt{~~~~thr\_join(tid, \&status);} \\
		\texttt{\}}
	\end{tabular}
	\end{center}
	\pause
	\linegap

@@

\begin{frame}{Previous Semesters}
	%% S'15, F'15, S'16: 48 groups signed up to use Landslide; 38 found bugs
	%% S'15, F'15, S'16, F'16, S'17: 79 groups signed up to use Landslide; 59 found bugs
	S'15, F'15, S'16, F'16, S'17, F'17: 94 groups signed up to use Landslide; 71 found bugs
	\linegap

	%%44 {\em deterministic} bugs (e.g. swexn, initialization)
	%% F17 lecture - stats up to S17
	%% 91 {\em deterministic} bugs (e.g. swexn, initialization)
	% S18 lecture - stats up to F17
	109 {\em deterministic} bugs (e.g. swexn, initialization)
	\linegap

@@

\begin{frame}{User Study}
	\textbf{Try Landslide on your P2!}
	\begin{itemize}
		\item Bare minimum effort: No more than 1 hour
		\begin{itemize}
			\item Clone a github URL, run setup script, run tests, answer survey
			\item Landslide will automatically report test results (as described below)
		\end{itemize}
		\item Full study plan: 4-8 hours of active attention
		\begin{itemize}
			\item (Estimated, including time to diagnose and fix bugs)
			\item However, many tests should run passively overnight -- start soon!
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	Prerequisites
	\begin{itemize}
		\item You {\em must} pass the P2 hurdle before using Landslide.
		\begin{itemize}
			\item {\tt startle}, {\tt agility\_drill}, {\tt cyclone}, {\tt join\_specific\_test}, {\tt thr\_exit\_join}
		\end{itemize}
		\item Recommended to attempt several stress tests, e.g.:
		\begin{itemize}
			\item {\tt juggle 4 3 2 0}, {\tt multitest}, {\tt racer}, {\tt paraguay}
		\end{itemize}

@@

\begin{frame}{User Study - Additional Information}
	Human Subjects Research
	\begin{itemize}
		\item CMU IRB has approved this study
		\item Landslide will collect results while you use it
		\begin{itemize}
			\item Record commands issued, take snapshots of your P2 code
			\item All data will be anonymized before publication
		\end{itemize}
		\item No coercion: {\bf \em There is no penalty for not participating.}
		\begin{itemize}
			\item I am not on course staff, cannot influence your grade
			\item Course staff will not have access to study data during semester
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	Risks \& Benefits
	\begin{itemize}
		\item Benefit: Landslide may help you find/fix bugs, improving your grade!
		\item Risk: Landslide may find no bugs and be a waste of your time.
		\item Benefit: You might learn something...
	\end{itemize}
	% Say: Accordingly, although I hope you will use landslide enough to get good results for me, you should also responsibly allocate time between testing with landslide versus actually implementing your P2.
\end{frame}

@@

	To participate\dots
	\begin{itemize}
		%\item Review this lecture and study information sheet
		\item Meet prerequisites of passing P2 tests
		\item Complete sign-up form online to get further instructions
			\begin{itemize}
				\item (watch your email for the link)
				%\item \url{http://tinyurl.com/landslide-p2-s17}
			\end{itemize}
		\item Optional ``Landslide clinic'' for in-person tech (or moral) support
			\begin{itemize}
				\item Next week, room and time TBD
			\end{itemize}
		%\item Use Landslide in addition to stress tests until P2 is due!
	\end{itemize}
	%% im out of time :<<
	%\linegap

@@

>>>>

\usepackage{manfnt}
\usepackage{wasysym}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{url}
\usepackage{ulem}
\usepackage{marvosym}
\usepackage{skull}
\usepackage{proof}
\usepackage{array}
\usepackage{colortbl}
\usepackage{xspace}
\setbeamertemplate{navigation symbols}{}

\newcommand\classname{CMPSC 473\xspace}
\title[Landslide]{{\bf Landslide: Concurrency Bug-Finding for Pebbles}}
\subtitle[]{ {\em more clever than } ``\texttt{agility\_drill}'' {\em since 2011.}}
\author[Ben Blum]{Ben Blum \texttt{(bblum@andrew.cmu.edu)}}

\institute[PSU \classname]{Penn State University - \classname}
\date[]{2018, April 3}

@@

\begin{document}
\renewcommand{\inserttotalframenumber}{42}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

\begin{frame}{Outline}
	\textbf{Theory: Seeing concurrency bugs in a new way}
	\begin{itemize}
		\item Examples
		\item Tabular execution traces
		\item The execution tree
	\end{itemize}
	{\bf Research Technique: ``Systematic testing''}
	% Say: A way of enumerating all possible interleavings of a concurrent test case.
	\begin{itemize}
		\item Preemption points
		\item Challenges and feasibility
	\end{itemize}
	{\bf Tool: Landslide}
	\begin{itemize}
		\item Automatically choosing preemption points
		\item Demo (the fun part)
		\item User study (that's you!)
			% !!! SAY !!! : If you are struggling, this is not for you, etc.
	\end{itemize}
\end{frame}

@@

% buggy atomics

\begin{frame}{Case Study 1}
	\begin{center}
			\fbox{
			\begin{tabular}{l}
				libatomic/atomic.S \\
				\hline
				\\
				\texttt{/* int atomic\_fetch\_add(volatile int* obj, int arg); */} \\
				\texttt{.global atomic\_fetch\_add} \\
				\texttt{atomic\_fetch\_add:} \\
				\texttt{~~mov 4(\%esp), \%ecx~~~~~~// obj} \\
				\texttt{~~mov 8(\%esp), \%edx~~~~~~// arg} \\
				\texttt{~~mov~~(\%ecx), \%eax~~~~~~// return old value of obj} \\
				\texttt{~~lock xadd \%edx, (\%ecx) // atomically increment obj} \\
				\texttt{~~ret} \\
			\end{tabular}
			}
		\pause
		\linegap

		\begin{itemize}
			\item Loading {\tt \%ecx} separately breaks atomicity!
			\item Two threads can see the same return value.
		\end{itemize}
	\end{center}
\end{frame}
\begin{frame}{Thread Interleavings (``good'' case)}
	\begin{center}
		\begin{tabular}{|l|l|}
			\hline
			\cellcolor{thread1} {\bf Thread 1} & \cellcolor{thread2} {\bf Thread 2} \\
			\hline
			\texttt{mov 4(\%esp), \%ecx} & \\
			\texttt{mov 8(\%esp), \%edx} & \\
			\texttt{mov~~(\%ecx), \%eax} & \\
			\texttt{lock xadd \%edx, (\%ecx)} & \\
			\texttt{ret} & \\
			& \texttt{mov 4(\%esp), \%ecx} \\
			& \texttt{mov 8(\%esp), \%edx} \\
			& \texttt{mov~~(\%ecx), \%eax} \\
			& \texttt{lock xadd \%edx, (\%ecx)} \\
			& \texttt{ret} \\
			\hline
		\end{tabular}
	\end{center}
\end{frame}
\begin{frame}{Thread Interleavings (different ``good'' case)}
	\begin{center}
		\begin{tabular}{|l|l|}
			\hline
			\cellcolor{thread1} {\bf Thread 1} & \cellcolor{thread2} {\bf Thread 2} \\
			\hline
			\texttt{mov 4(\%esp), \%ecx} & \\
			\texttt{mov 8(\%esp), \%edx} & \\
			& \texttt{mov 4(\%esp), \%ecx} \\
			& \texttt{mov 8(\%esp), \%edx} \\
			& \texttt{mov~~(\%ecx), \%eax} \\
			& \texttt{lock xadd \%edx, (\%ecx)} \\
			& \texttt{ret} \\
			\texttt{mov~~(\%ecx), \%eax} & \\
			\texttt{lock xadd \%edx, (\%ecx)} & \\
			\texttt{ret} & \\
			\hline
		\end{tabular}
	\end{center}
\end{frame}
\begin{frame}{Thread Interleavings (race condition)}
	\begin{center}
		\begin{tabular}{|l|l|}
			\hline
			\cellcolor{thread1} {\bf Thread 1} & \cellcolor{thread2} {\bf Thread 2} \\
			\hline
			\texttt{mov 4(\%esp), \%ecx} & \\
			\texttt{mov 8(\%esp), \%edx} & \\
			\texttt{mov~~(\%ecx), \%eax} & \\
			& \texttt{mov 4(\%esp), \%ecx} \\
			& \texttt{mov 8(\%esp), \%edx} \\
			& \texttt{mov~~(\%ecx), \%eax} \\
			& \texttt{lock xadd \%edx, (\%ecx)} \\
			& \texttt{ret} \\
			\texttt{lock xadd \%edx, (\%ecx)} & \\
			\texttt{ret} & \\
			\hline
		\end{tabular}
	\end{center}
\end{frame}

% paradise lost

\begin{frame}{Case Study 2}
	\begin{center}
		\begin{tabular}{llll}
			\fbox{
			\begin{tabular}{l}
				Consumer thread \\
				\hline
				\\
				\texttt{mutex\_lock(mx);} \\
				\\
				\texttt{if~(!work\_exists())} \\
				\texttt{~~~~cond\_wait(cvar,~mx);} \\
				\texttt{work~=~dequeue();} \\
				\\
				\texttt{mutex\_unlock(mx);} \\
				\texttt{access(work->data);} \\
			\end{tabular}
			}
			& & &
			\fbox{
			\begin{tabular}{l}
				Producer thread \\
				\hline
				\\
				\texttt{mutex\_lock(mx);} \\
				\\
				\texttt{enqueue(work);} \\
				\texttt{signal(cvar);} \\
				\\
				\texttt{mutex\_unlock(mx);} \\
				\\
				\\
			\end{tabular}
			}
		\end{tabular}
		\pause
		\linegap

		\begin{itemize}
			\item See {\bf Synchronization} lecture!
			\item {\tt if} vs {\tt while}: Two consumers can race to make one fail.
		\end{itemize}
	\end{center}
\end{frame}
\begin{frame}{Thread Interleavings (``good'' case)}
	\begin{center}
		\begin{tabular}{|l|l|l|}
			\hline
			\cellcolor{thread1} {\bf Thread 1} & \cellcolor{thread2} {\bf Thread 2} & \cellcolor{thread3} {\bf Thread 3} \\
			\hline
			\small \texttt{lock(mx);} & & \\
			\small \texttt{if~(!work\_exists())} & & \\
			\small \texttt{~~~~wait(cvar,~mx);} & & \\

@@

\begin{frame}{Testing}
	How can programmers be confident in the correctness of their code?
	\begin{itemize}
		\item Unit tests
		\begin{itemize}
			\item good for basic functionality, bad for concurrency
		\end{itemize}
		\item Stress tests
		\begin{itemize}
			\item state of the art in \classname
		\end{itemize}
		\item Theorem proving
		\begin{itemize}
			\item heavy burden on the programmers
		\end{itemize}
		\item \sout{Releasing to paying customers and worrying about correctness later}
			% Say... "and I'm hoping to prevent you from finding yourself in a future where this happens!"
	\end{itemize}
	\linegap

@@

\begin{frame}{Preemption Point Example ({\tt atomic\_fetch\_add()})}
	\fbox{
	\begin{tabular}{rll}
		\multicolumn{3}{l}{\texttt{/* int atomic\_fetch\_add(volatile int* obj, int arg); */}} \\
		1 & \texttt{.global atomic\_fetch\_add} \\
		2 & \texttt{atomic\_fetch\_add:} \\
		3 & \texttt{~~mov 4(\%esp), \%ecx} \\
		  & & \footnotesize (preemption point A) \\
		4 & \texttt{~~mov 8(\%esp), \%edx} \\
		  & & \footnotesize (preemption point B) \\
		5 & \texttt{~~mov~~(\%ecx), \%eax} \\
		  & & \footnotesize (preemption point C) \\
		6 & \texttt{~~lock xadd \%edx, (\%ecx)} \\
		  & & \footnotesize (preemption point D) \\
		7 & \texttt{~~ret} \\
	\end{tabular}
	}
	\linegap

@@

\begin{frame}{Preemption Point Example ({\tt atomic\_fetch\_add()})}
	\fbox{
	\begin{tabular}{rll}
		\multicolumn{3}{l}{\texttt{/* int atomic\_fetch\_add(volatile int* obj, int arg); */}} \\
		1 & \texttt{.global atomic\_fetch\_add} \\
		2 & \texttt{atomic\_fetch\_add:} \\
		3 & \texttt{~~mov 4(\%esp), \%ecx} \\
		  & & \footnotesize (preemption point A) \\
		4 & \texttt{~~mov 8(\%esp), \%edx} \\
		  & & \footnotesize (preemption point B) \\
		5 & \texttt{~~mov~~(\%ecx), \%eax} \\
		  & & \footnotesize (preemption point C) \\
		6 & \texttt{~~lock xadd \%edx, (\%ecx)} \\
		  & & \footnotesize (preemption point D) \\
		7 & \texttt{~~ret} \\
	\end{tabular}
	}
	\linegap

	Here, only preemption point B will break atomicity. {\bf ~}\\
	All other interleavings are benign.{\bf ~}

@@

\begin{frame}{About The Project}
	\textbf{About me}: Final year graduate student at Carnegie Mellon
	\begin{itemize}
		\item TAed CMU OS (same thread project) for 3 semesters %during undergrad
		%\item 1st graduate year was 5th year M.S., rest in Ph.D. program
		\item Landslide's publication history
		\begin{itemize}
			\item Master's thesis (CMU)
			\begin{itemize}
				\item \url{http://www.contrib.andrew.cmu.edu/~bblum/thesis.pdf}
			\end{itemize}
			\item Conference paper (OOPSLA 2016)
			\begin{itemize}
				\item \url{http://www.contrib.andrew.cmu.edu/~bblum/oopsla.pdf}
			\end{itemize}
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	{\bf About Landslide}
	\begin{itemize}
		\item Runs your code in a simulator (Bochs) which traces:
			\begin{itemize}
				\item Every instruction executed
				\item Every memory access read/written
			\end{itemize}
			% FIXME for next year -- how much slowdown is there? 20x?
		%\item Originally supported only P3s; can now test P2s fully-automated
		\item {\em Landslide} shows how your {\em Pebbles} programs may not be stable.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Big Picture: Execution Tree Exploration}
	\textbf{Backtracking}
	\begin{itemize}
		\item Each time test completes, identify a PP to replay differently
		\item Reset thread library state and start over
		% \item Implemented with fork() in Bochs
		\item Replay test from the beginning, with a different interleaving
	\end{itemize}
	\pause
	\linegap

	{\bf Controlling scheduling decisions}
	\begin{itemize}
		\item Tool must control all sources of nondeterminism
		\item For thread libraries, just timer interrupts
		%\item In 15-410, just timer and keyboard interrupts
			% optional - say: "in real kernels, disk/network io interrupts"
		\item Landslide repeatedly fires timer ticks until desired thread is run.
	\end{itemize}
\end{frame}

%\begin{frame}{Landslide \& You}
%	\begin{center}
%	\includegraphics[width=0.85\textwidth]{landslide-new-blank.pdf}
%	\end{center}
%\end{frame}

@@

\begin{frame}{Identifying Bugs}
	Landslide can {\em definitely discover:}
	\begin{itemize}
		\item Assertion failures % Say: note: the better your asserts are..!
		\item Segfaults
		\item Deadlock
		\item Use-after-free / double-free
	\end{itemize}
	\linegap
	Landslide can {\em reasonably suspect:}
	\begin{itemize}
		\item Infinite loop (see: halting problem)
		\item Data race bugs
	\end{itemize}
\end{frame}

@@

	Data races are {\em not necessarily} bugs, just highly suspicious!
	\begin{itemize}
		%\item Bakery alg: Is {\tt number[i]=max(number[0],number[1])+1} bad?
		\item Atomic load/store: Is {\tt mov \%ecx, \%edx} bad?
		%\item Is unprotected {\tt foobar\_statistics\_counter++} bad?
		\item What about unprotected {\tt next\_thread\_id++}?
		\item ``If threads interleaved the wrong way here, it {\em might} crash later.''
		\begin{itemize}
			\item Hmmm...
		\end{itemize}
	\end{itemize}
\end{frame}

@@

	State of the art tools choose a fixed set of preemption points.
	\begin{itemize}
		\item E.g., ``all thread API calls'' or ``all mutex locks/unlocks''
		\item Depending on length of test, completion time is unpredictable.
		\item More often, a subset is better in terms of time/coverage. 
	\end{itemize}
	\pause
	\linegap
	Current systematic testing model is not user-friendly.
	\begin{itemize}
		\item Tool: ``I want to use these PPs, but can't predict completion time.''
		\item User: ``I have 16 CPUs and 24 hours to test my program.''
	\end{itemize}
	\linegap
	{\em Stress testing allows user to choose total run time -- can we offer this too?}
\end{frame}

@@

\begin{frame}{Test Suite}
	Landslide can check the following thread library tests:
	%Landslide ships with 6 approved test cases:
	%\linegap

	Atomic tests
	\begin{itemize}
		\item {\tt atomic\_fetch\_add}
		\item (etc...)
	\end{itemize}

	Locking tests
	\begin{itemize}
		\item mutex: {\tt mutex\_test}
		\item cvars: {\tt broadcast\_test}, {\tt paraguay}
		\item semaphores: {\tt paradise\_lost}
		\item rwlocks: {\tt rwlock\_downgrade\_read\_test}
	\end{itemize}

	Threading tests
	\begin{itemize}
		\item {\tt thr\_exit\_join}
	\end{itemize}
\end{frame}

@@

	Sample code from {\tt cyclone}: % abbreviated
	\begin{center}
	\begin{tabular}{l}
		\texttt{for (i = 0; i < 64; i++) \{} \\
		%\texttt{for (i = 0; i < MAX\_MISBEHAVE /* 64 */; i++) \{} \\
		%\texttt{~~~~misbehave(i);} \\
		\texttt{~~~~tid = thr\_create(child\_fn, i);} \\
		\texttt{~~~~thr\_join(tid, \&status);} \\
		\texttt{\}}
	\end{tabular}
	\end{center}
	\pause
	\linegap

@@

\begin{frame}{Previous Semesters in CMU's OS class}
	%% S'15, F'15, S'16: 48 groups signed up to use Landslide; 38 found bugs
	%% S'15, F'15, S'16, F'16, S'17: 79 groups signed up to use Landslide; 59 found bugs
	S'15, F'15, S'16, F'16, S'17, F'17: 94 groups signed up to use Landslide; 71 found bugs
	\linegap

	%%44 {\em deterministic} bugs (e.g. swexn, initialization)
	%% F17 lecture - stats up to S17
	%% 91 {\em deterministic} bugs (e.g. swexn, initialization)
	% S18 lecture - stats up to F17

	%109 {\em deterministic} bugs (e.g. swexn, initialization)
	%\linegap

@@

\begin{frame}{User Study}
	\textbf{Try Landslide on your project!}
	\begin{itemize}
		\item Bare minimum effort: No more than 1 hour
		\begin{itemize}
			\item Clone a github URL, run setup script, run tests, answer survey
			%\item Landslide will automatically report test results (as described below)
		\end{itemize}
		\item Full study plan: 4-8 hours of active attention
		\begin{itemize}
			\item (Estimated, including time to diagnose and fix bugs)
			\item However, many tests should run passively overnight -- start soon!
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	Prerequisites
	\begin{itemize}
		\item Atomic tests: Landslide requires working {\tt thr\_create()}.
		\begin{itemize}
			\item You must first pass {\tt startle}
		\end{itemize}
		\item Locking and threading tests: also need {\tt thr\_exit()} and {\tt thr\_join()}
		\begin{itemize}
			\item Must pass {\tt agility\_drill}, {\tt cyclone}, {\tt join\_specific\_test}, {\tt thr\_exit\_join}
		\end{itemize}
		%\item You {\em must} pass the P2 hurdle before using Landslide.
		%\begin{itemize}
		%	\item {\tt startle}, {\tt agility\_drill}, {\tt cyclone}, {\tt join\_specific\_test}, {\tt thr\_exit\_join}
		%\end{itemize}
		\item Recommended to attempt several stress tests, e.g.:
		\begin{itemize}
			\item {\tt juggle 4 3 2 0}, {\tt multitest}, {\tt racer}, {\tt paraguay}
		\end{itemize}

@@

\begin{frame}{User Study - Additional Information}
	Human Subjects Research
	\begin{itemize}
		\item CMU IRB has approved this study
		\item I will survey you when the project is done
		\begin{itemize}
			\item How many bugs did you find, was it easy to use...
			\item All survey responses will be anonymized before publication
		\end{itemize}
		%\item Landslide will collect results while you use it
		%\begin{itemize}
		%	\item Record commands issued, take snapshots of your P2 code
		%	\item All data will be anonymized before publication
		%\end{itemize}
		\item No coercion: {\em There is no penalty for not participating.}
		\begin{itemize}
			\item I am not on course staff, cannot influence your grade
			\item Course staff will not have access to study data during semester
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	Risks \& Benefits
	\begin{itemize}
		\item Benefit: Landslide may help you find/fix bugs, improving your grade!
		\item Risk: Landslide may find no bugs and be a waste of your time.
		% sorry dave
		%\item Benefit: You might learn something...
	\end{itemize}
	% Say: Accordingly, although I hope you will use landslide enough to get good results for me, you should also responsibly allocate time between testing with landslide versus actually implementing your P2.
\end{frame}

@@

	To participate\dots
	\begin{itemize}
		%\item Review this lecture and study information sheet
		\item Meet prerequisites of passing thread library tests
		\item Complete sign-up form online to get further instructions
			\begin{itemize}
				\item (link will be emailed to you soon)
				%\item \url{http://tinyurl.com/landslide-p2-s17}
			\end{itemize}
		\item Email me with any Landslide-related questions: {\sf bblum@cs.cmu.edu}
			\begin{itemize}
				\item OK: ``How do I interpret the html output as a thread interleaving?''
				\item OK: ``Landslide said it crashed and to email you for tech support!''
				\item NOT: ``Landslide found a bug, what's wrong with my code?''
				\item NOT: ``Landslide found a bug, how do I fix it?''
			\end{itemize}
		%\item Optional ``Landslide clinic'' for in-person tech (or moral) support
		%	\begin{itemize}
		%		\item Next week, room and time TBD
		%	\end{itemize}
		%\item Use Landslide in addition to stress tests until P2 is due!
	\end{itemize}
	%% im out of time :<<
	%\linegap

@@

\breakslide{\Huge Bonus Slides}
